---
title: "Lab week 6"
author: "Lucas Boyd"
date: "11/1/2021"
output: html_document
---

```{r setup, include=TRUE, message = FALSE, warning = FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
library(here)
library(tidyverse)
library(palmerpenguins)
library(broom)
library(equatiomatic)
```

## example of a rank-based test

We'll make our own samples, using a pseudo-random generator.

```{r}
set.seed(1414)
# create a sample of integer values (max value, number of values, replace is whether numbers can repeat)
gp_1 <- sample.int(20, size = 15, replace = TRUE) # create a sample of integer values (max value, number of values, replace is whether numbers can repeat)

set.seed(1424)
gp_2 <- sample.int(30, size = 15, replace = TRUE)
```
Quick histograms for exploratory purposes.
```{r}
hist(gp_1)

hist(gp_2)

```


Try a t-test. 

```{r}
t.test(gp_1, gp_2)
```

This p-value means if these samples were drawn from populations with the same mean, there is a 19.8 percent chance that we would have found a difference of sample means at least this large by random chance. 

Retain (fail to reject) the null hypothesis! There is no significant difference in means between group 1 and group 2. 

Now let's compare this outcome with a rank-based test.

## Mann Whitney U unpaired rank-based test

```{r}
mwu <- wilcox.test(gp_1, gp_2)

mwu
```

The p value means that 

if these samples were drawn from populations with the same ranks (medians), the probability of finding two samples with ranks at least as different as those in our samples is 28%.

There is no significant different in ranks between group 1 and group 2 (statistical summary).

Median scores for group 1 (M = 14) and group 2 (M = 12) did not differ signficantly (U(df) = 86, p = 0.28).

kruskal.test() for rank based test with more than 2 groups

## Linear regression
Simple linear regression (single dependent variable, a single independent variable)

```{r}
# make an exploratory plot of penguin body mass (y) vs. flipper length (x)
ggplot(data = penguins, aes(x = flipper_length_mm, y = body_mass_g)) +
  geom_point() +
  geom_smooth(method = "lm") # adding a trendline
```

Find a linear regression model using ordinary least squares describing the relationship between flipper length and body mass for these penguins. 

3 pieces:
- what type of model?
- what is the relationship to model (Dependent variable (y) as a function of Independent variable (x)
- where is the data that's used to create this model?

lm() for simple linear models

```{r}
penguin_lm <- lm(body_mass_g ~ flipper_length_mm, data = penguins)
```

y = mx + b 
Body mass = 49.7(flipper_length) - 5780.8

Which means I expect a 1mm increase in flipper length to be associated with a 49.7g increase in body mass.

penguin_lm$coefficients[2] to pull out a specific value from the vector

### broom package returns model outputs as tidy data frames

```{r}
penguin_lm_tidy <- tidy(penguin_lm)
# glance gets you model-wide measures of uncertainty, to be continued...
glance(penguin_lm)
```

How can I actually include my model equation in a report?

```{r}
extract_eq(model = penguin_lm, use_coefs = TRUE) # this makes a nice knitted lineer model equation that is dank and reproducible
```

plot() function to visualize linear models
```{r}
plot(penguin_lm)
```



