---
title: "Lab week 6"
author: "Lucas Boyd"
date: "11/1/2021"
output: html_document
---

```{r setup, include=TRUE, message = FALSE, warning = FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
library(here)
library(tidyverse)
library(palmerpenguins)
library(broom)
library(equatiomatic)
```

## example of a rank-based test

We'll make our own samples, using a pseudo-random generator.

```{r}
set.seed(1414)
# create a sample of integer values (max value, number of values, replace is whether numbers can repeat)
gp_1 <- sample.int(20, size = 15, replace = TRUE) # create a sample of integer values (max value, number of values, replace is whether numbers can repeat)

set.seed(1424)
gp_2 <- sample.int(30, size = 15, replace = TRUE)
```
Quick histograms for exploratory purposes.
```{r}
hist(gp_1)

hist(gp_2)

```


Try a t-test. 

```{r}
t.test(gp_1, gp_2)
```

This p-value means if these samples were drawn from populations with the same mean, there is a 19.8 percent chance that we would have found a difference of sample means at least this large by random chance. 

Retain (fail to reject) the null hypothesis! There is no significant difference in means between group 1 and group 2. 

Now let's compare this outcome with a rank-based test.

## Mann Whitney U unpaired rank-based test

```{r}
mwu <- wilcox.test(gp_1, gp_2)

mwu
```

The p value means that 

if these samples were drawn from populations with the same ranks (medians), the probability of finding two samples with ranks at least as different as those in our samples is 28%.

There is no significant different in ranks between group 1 and group 2 (statistical summary).

Median scores for group 1 (M = 14) and group 2 (M = 12) did not differ signficantly (U(df) = 86, p = 0.28).

```{r}

```

